---
output:
  pdf_document:
     includes:
       in_header: finance.sty
---

```{r setup, include=FALSE}
library(tidyverse)
library(wrapr)
library(stringr)
library(lubridate)
inds <- read.csv("Industry49_data.csv", header = T)

# Function to calculate efficient frontier
efficient_front <- function(M = 30, assets = c(), N = 0,
                            risk_free = F){
  
  # This takes either a vector of assets, or a number (N) 
  # of assets to choose at random
  
  # Get number of assets if supplied, asset names if not
  if (length(assets) == 0){
    industries = names(R_bar)[names(R_bar) != "Market"]
    assets = sample(industries, N)
  } else {
    N = length(assets)
  }
  
  # Get the necessary matrices/vectors
  mus = seq(-0.05, 0.20, length.out = M) # return values 
  V = v_mat[assets, assets] # variance-covar
  V_inv = solve(V)
  R = R_bar[assets] # expected returns on assets
  one = rep(1.0, N)
  
  # Calculate efficient frontier portfolio weights
  if (risk_free){
    Rf = 0.01 # default risk-free rate is 1%
    Rf_1 = one * Rf
    
    opt_w_rf <- function(mu){
      w_opt = as.numeric((mu - Rf) / (t(R - Rf_1) %*% V_inv %*% (R - Rf_1))) * 
        V_inv %*% (R - Rf_1)
    }
    
    eff_front = t(sapply(mus, opt_w_rf)) # get optimal w
    colnames(eff_front) = assets
    
  } else {
    # Get the optimal portfolios
    w_r = V_inv %*% R / as.numeric(one %*% V_inv %*% R)
    w_1 = V_inv %*% one / as.numeric(one %*% V_inv %*% one)
    
    # Calculate alpha, as function of mu
    get_alpha = function(mu){
      alpha = as.numeric(mu - t(w_1) %*% R) / 
        as.numeric(t(w_r) %*% R - t(w_1) %*% R)
    }
    
    get_alpha = Vectorize(get_alpha)
    
    # Calculate efficient frontier: optimal portfolios at each return
    alphas = get_alpha(mus)
    
    wts_1 = alphas %*% t(w_r)
    wts_2 = (1 - alphas) %*% t(w_1)
    eff_front = wts_1 + wts_2
  }
  
  port_vars = vector("numeric")
  for (i in 1:M){
    v = t(eff_front[i, ]) %*% V %*% eff_front[i, ]
    port_vars = append(port_vars, v)
  }
  
  eff_front = cbind.data.frame(mus, eff_front)
  eff_front[["p_sd"]] = sqrt(port_vars)
  
  indiv_assets = data.frame(
    indust = assets,
    eR = R, 
    var = sqrt(diag(V))
  )
  
  # Return args
  return(list(eff_front = eff_front, indiv_assets = indiv_assets))
  
}
```

\begin{flushleft}
	Nick Hoffman \\
	Finance I, Spring 2020 A \\
	Assignment 1 \\
\end{flushleft}

\begin{enumerate}
	\item Useful facts
	\begin{enumerate}
		\item $ \ev[\w ^\p R ] = \w^\p \bar{R} $, where $\bar{R}$ is the $ N\times 1 $ vector of expected returns.
		\[\ev[\w^\p \bar{R}] = \ev\left[\sum_{i = 1}^{N}\tilde{R}_n \w_n\right] = \sum_{i = 1}^N \w_n \ev[\tilde{R}_n] = \w^\p \bar{R} \]
		
		\item $ \frac{\partial \ev[R^\p \w]}{\partial \w_n} = R_n $, or the entire vector $  \frac{\partial \ev[R^\p \w]}{\partial \w} = R $
		
		As in (a),
		\[\ev[\w^\p \bar{R}] =\sum_{i = 1}^N \w_n \ev[\tilde{R}_n]\]
		and thus
		\[\frac{\partial \ev[R^\p \w]}{\partial \w_n} = \ev[\tilde{R}_n] = \bar{R}\]
		and so in vector form,
		\[ \frac{\partial \ev[R^\p \w]}{\partial \w} = R\]
		
		\item $ V(R^\p \w) = cov(R^\p \w, R^\p \w) $ where $ V $ is the variance-covariance matrix of returns, with typical element $ \sigma_{i,j} = cov(\tilde{R}_i, \tilde{R}_j) $.
		
		By definition, 
		\begin{align*}
		cov(R^\p \w, R^\p \w) &= \ev\left[\sum_{i = 1}^N \w_n \tilde{R}_n \cdot \sum_{i = 1}^N \w_n \tilde{R}_n\right] - \ev\left[\sum_{i = 1}^N \w_n \tilde{R}_n\right]^2 \\
		&= \sum_{i = 1}^N \sum_{j = 1}^N \w_i \w_j \ev[R_i R_j] - \sum_{i = 1}^N \sum_{j = 1}^N \w_i \w_j \ev[R_i] \ev[R_j] \\
		&= \sum_{i = 1}^N \sum_{j = 1}^N \w_i\w_j cov(R_i, R_j) \\
		&= \w^\p V\w
		\end{align*}
		
		\item $ cov(R^\p \w, R_n) = \w^\p V e_n $, where $ e_n = [0, \dots, 0, 1, 0, \dots, 0]^\p $, with 1 in the $ n^{th} $ position.
		
		Using the same definition for the covariance as above, 
		\begin{align*}
		cov(R^\p \w, R_n) &= \ev[R^\p \w R_n] - \ev[R^\p \w]\ev[R_n] \\
		&= \sum_{i = 1}^N \w_i \big(\ev[R_i R_n] - \ev[R_i]\ev[R_n]\big) \\
		&= \sum_{i = 1}^N \w_i cov(R_i, R_n)
		\end{align*} 
		The final sum above is the $ n^{th} $ entry of the matrix $ \w^\p V $, and thus $ cov(R^\p \w, R_n) = \w^\p V e_n $
		
		\item $ \frac{\partial V(R^\p \w)}{\partial \w_n} = 2 cov(R^\p \w, R_n) $
		
		As in part (c), 
		\[V(R^\p \w) = \sum_{i = 1}^N \sum_{j = 1}^N \w_i \w_j cov(R_i, R_j)\]
		Thus,

		\[\frac{\partial V(R^\p \w)}{\partial \w_n} = 2 \sum_{i = 1}^N \w_i cov(R_i, R_n) = 2 cov(R^\p \w, R_n) \]
		where the final equality follows from (d).
	\end{enumerate}
	\item The portfolio problem with risky assets is 
	\begin{gather*}
	\min_\w \frac{1}{2}\w^\p V \w \\
	\text{s.t.} \\
	\w^\p \bar{R} \geq \mu \\
	\w^\p \one = 1
	\end{gather*}
	where $ \one $ is the $N$-vector with 1 in every element. The Lagrangean for this problem is
	\[\lagr = \frac{1}{2}\wp V \w + \lambda_1\big[\mu - \wp\bar{R}\big] + \lambda_2\big[1 - \wp \one\big]\]
	
	with first-order conditions
	\begin{align}
	V\w &= \lambda_1\bar{R} - \lambda_2\one \label{foc_1_1} \\
	\mu &= \wp \bar{R} \label{foc_1_2} \\
	\wp \one &= 1 \label{foc_1_3}
	\end{align}
	To begin, premultiply (\ref{foc_1_1}) by $ \inv{V} $, yielding
	\[\w = \lambda_1 \inv{V}\bar{R} + \lambda_2\inv{V}\one\]
	Now, define the following two efficient portfolios:
	\begin{align*}
	\w_R &= \frac{\inv{V}\bar{R}}{\one^\p \inv{V} \bar{R}} \\
	\w_1 &= \frac{\inv{V} \one}{\one^\p \inv{V} \one}
	\end{align*}
	Thus, the form for $ \w $ above can be transformed in the following way:
	\begin{align*}
	\w^* &= \lambda_1 \inv{V}\bar{R} + \lambda_2\inv{V}\one \\ 
	&= \lambda_1\left(\frac{\one^\p \inv{V}\bar{R}}{\one^\p \inv{V}\bar{R}}\right)\inv{V}\bar{R} + \lambda_2\left(\frac{\one^\p \inv{V}\one}{\one^\p \inv{V}\one}\right)\inv{V}\one \\
	&= \big(\lambda_1\one^\p \inv{V}\bar{R}\big)\w_R + \big(\lambda_2\one^\p \inv{V}\one\big)\w_1 \\
	&\equiv \alpha \w_R + (1 - \alpha)\w_1
	\end{align*}
	
	To show the final equality (i.e., that the weights on portfolios $ \w_R $ and $ w_1 $ sum to unity), we premultiply both sides by $ \one^\p $ and impose  (\ref{foc_1_3}):
	\[\onep \w^*= \big(\lambda_1\one^\p \inv{V}\bar{R}\big)\onep w_R + \big(\lambda_2\one^\p \inv{V}\one\big)\onep w_1 \]
	Because $ \w^* $, $ \w_R $, and $ \w_1 $ are portfolios, the above implies that 
	\[1 = \big(\lambda_1\one^\p \inv{V}\bar{R}\big) + \big(\lambda_2\one^\p \inv{V}\one\big)\]
	To solve for $\alpha$, we impose (\ref{foc_1_2}):
	\begin{align*}
	\w^{*\p}\bar{R} = \mu &\implies \alpha \w_R^\p \bar{R} + (1 - \alpha) \w_1^\p \bar{R} = \mu \\
	& \implies \alpha = \frac{\mu - \w_1^\p \bar{R}}{\w_R^\p \bar{R} - \w_1^\p \bar{R}}
	\end{align*}
	
	\item To find the Global Minimum Variance (GMV) portfolio, the problem is similar:
	\begin{gather*}
	\min_\w \frac{1}{2}\w^\p V \w \\
	\text{s.t.} \\
	\w^\p \one = 1
	\end{gather*}
	
	The Lagrangean for this problem is 
	\[\lagr = \frac{1}{2} \wp V \w + \lambda[1 - \wp \one ]\]
	From the first-order condition for $\omega$, we can use the same method as in question 2:
	\begin{align*}
	\w^* &= \lambda \inv{V} \one  \\
	&= \lambda\left(\frac{\one^\p \inv{V}\one}{\one^\p \inv{V}\one}\right)\inv{V}\one \\
	&= \big(\lambda \onep \inv{V}\one \big)\w_1
	\end{align*}
	Because $ \w $ and $ \w_1 $ are both portfolios, premultiplying the final line above by $ \onep $ shows that $ (\lambda \onep \inv{V}\one) = 1 $. Thus, $ \w_1 $ is the GMV portfolio
	
	\item With both risky assets and one risk-free asset, the problem becomes
	\begin{gather*}
	\min_\w \frac{1}{2}\w^\p V \w \\
	\text{s.t.} \\
	(1 - \w^\p \one)R_f + \w^\p \bar{R} \geq \mu
	\end{gather*}
	\begin{enumerate}
		\item The Lagrangian for this problem is as follows:
		\[\lagr = \frac{1}{2}\w^\p V \w + \lambda \big[\mu - (1 - \w^\p \one)R_f - \w^\p \bar{R}\big] \]
		with first-order conditions
		\begin{align}
		V\w + \lambda(\one R_f - \bar{R}) &= \mu \label{foc_4_1}\\
		(1 - \w^\p \one)R_f + \w^\p \bar{R} &\geq \mu \label{foc_4_2}
		\end{align}
		To begin, premultiply (\ref{foc_4_1}) by $ \inv{V} $ (assuming that $ V $ is invertible):

		\begin{equation}\label{step2_4}
		\w = \lambda \inv{V} \big(\bar{R} - \one R_f\big)
		\end{equation}
		
		Similarly, from (\ref{foc_4_2}), 
		\[\big(\bar{R} - \one R_f\big)\wp = \mu - R_f\]
		Thus, premultiplying (\ref{step2_4}) by $ (\bar{R} - \one R_f) $ gives
		\[\lambda\big(\bar{R} - \one R_f\big) \inv{V} \big(\bar{R} - \one R_f\big)\]
		and thus
		\[\lambda = \frac{\mu - R_f}{(\bar{R} - \one R_f\big) \inv{V} \big(\bar{R} - \one R_f )}\]
		Therefore, the optimal $ \w $ is given by 
		\[\w = \frac{\mu - R_f}{(\bar{R} - \one R_f\big) \inv{V} \big(\bar{R} - \one R_f )} \inv{V} \big(\bar{R} - \one R_f\big) \]
		
		\item To verify that the efficient frontier is linear in $ \sigma_p $, note that the variance of this portfolio is given by 
		\begin{align*}
		\wp V\w &= \frac{\mu - R_f}{(\bar{R} - \one R_f\big) \inv{V} \big(\bar{R} - \one R_f )} \big(\bar{R} - \one R_f\big)^\p \inv{V} V \frac{\mu - R_f}{(\bar{R} - \one R_f\big) \inv{V} \big(\bar{R} - \one R_f )} \inv{V} \big(\bar{R} - \one R_f\big) \\
		&= \frac{(\mu - R_f)^2}{\inv{V} \big(\bar{R} - \one R_f\big)}
		\end{align*}
		And thus the variance is 
		\[\sigma_p = \sqrt{\wp V \w} = \frac{|\mu - R_f}{\sqrt{\inv{V} \big(\bar{R} - \one R_f\big)}}\]
		Thus, solving for the return $ \mu_p $ gives
		\[\mu_p = R_f \pm \sigma_p \sqrt{\inv{V} \big(\bar{R} - \one R_f\big)}\]
		Thus, the portfolio return is linear in the variance, and thus the efficient frontier is linear. 
		
		\item The optimal Sharpe Ratio portfolio is the portfolio $ \w $ which solves
		\[\max_\w \frac{\wp\bar{R} - R_f}{\sqrt{\wp V \w}}\]
		The first-order condition for this problem is 
		\[\frac{d}{d\w} = -\frac{1}{2}(\wp \bar{R}) - R_f (\wp V \w)^{-\frac{3}{2}} V\w + \bar{R}(\wp V \w)^{-\frac{1}{2}} = 0 \]
		
		The portfolio solving this problem is 
		\[\w^* = \frac{\inv{V}\big(\bar{R} - \one R_f\big)}{\onep \inv{V} \big(\bar{R} - \one R_f\big)} \]
		
		This is the portfolio where, given $ N $ risky assets, the efficient frontier for the assets is tangent to the efficient frontier for the risky assets plus a risk free asset. Thus, this portfolio is also the efficient portfolio that includes only risky assets, when both risky and risk-free assets are available. 
	\end{enumerate}
	
```{r q5, include=FALSE}
# Clean the excess returns data
ers <- inds %>% 
  mutate(date = paste(year, month, day, sep = "-"),
         date = ymd(date)) %>% 
  select(date, starts_with("eR")) %>% 
  pivot_longer(-date, names_to = "industry", values_to = "eR") %>% 
  mutate(industry = str_remove_all(industry, "eR_"),
         industry = str_remove_all(industry, "I_"))

# a) Expected return by industry
exp_ers <- ers %>% 
  group_by(industry) %>% 
  summarise(
    r_bar = mean(eR, na.rm = T),
    r_bar = r_bar * 12 + 0.01 # convert to ann, add RfR
  )

R_bar <- exp_ers[["r_bar"]] # get vector, which is what we're after
names(R_bar) <- exp_ers[["industry"]] # want to keep track

# b) Variance-covariance matrix
v_mat <- ers %>% 
  pivot_wider(names_from = industry, values_from = eR) %>% 
  select(-date) %>% 
  as.matrix() %>% 
  cov(use = "complete.obs")

v_mat <- v_mat * 12 # convert to annual

# c) Sharpe ratio
sds <- sqrt(diag(v_mat)) # standard deviations
sharpe <-( R_bar - 0.01) / sds

returns_data <- cbind.data.frame(
  Industry = names(R_bar), 
  excess_r = R_bar * 100, 
  var = sqrt(diag(v_mat)) * 100,
  stand_dev = sds * 100, 
  sharpe = sharpe * 100
)

```

  \item 
```{r q5a, echo=FALSE}
returns_data %>% 
  knitr::kable(caption = "Returns data")
```
  
  \item
  \item 
  \item
\end{enumerate}